\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-03-24 redaction/dhp_sept2016.tex
Se positionner par rapport à Hummels sur les effets de composition

Before making any definite statement about this, it is worth emphasizing that the evolution of transport costs depends on the one hand of the evolution of per product and per partner costs and on the other hand of the evolution of the composition of trade.
Total transport costs may have decreased over time because the share of neighbour countries in total US trade or the share of cheaper to transport goods have increased, independently of any change in transport costs \textit{per se}. As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs. To do this, we estimate the following equations for observed costs and the estimated ad-valorem component using OLS:
%the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}$ and $$$%\widehat{t}_{ikt}$ that have been previously estimated:
\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated both for observed costs and the estimated ad-valorem componen we write it only for  $\widehat{\tau}_{iskt}$ for sake of notational simplicity. For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
%$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}

>>>>

Before making any definite statement about this, it is worth emphasizing that the evolution of transport costs depends on the one hand of the evolution of per product and per partner costs and on the other hand of the evolution of the composition of trade.
Total transport costs may have decreased over time because the share of neighbour countries in total US trade or the share of cheaper to transport goods have increased, independently of any change in transport costs \textit{per se}. As argued by \citet{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of per product- and per partner- transport costs.

\textbf{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}

\textbf{argument general}
In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.\footnote{In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}

\textbf{argument sur les effets de composition}
If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146.} By contrast, we provide an explicit decomposition of this additive cost component over time: By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. As well, we are able to do so for the multiplicative component of trade costs. We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects. This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}). Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.

\textbf{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps. Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes. Ligne 4, a faire, la somme des deux, estimee? Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}

Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy. Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components. \textbf{Pertinence de cet argument? A reformuler.}

To do this, we estimate the following equations for observed costs and the estimated ad-valorem component using OLS:
%the ad-valorem cost, based on the series obtained $\widehat{\tau}_{ikt}$ and $$$%\widehat{t}_{ikt}$ that have been previously estimated:
\footnote{Even if Equation (\ref{eq:compeffects_mult}) is estimated both for observed costs and the estimated ad-valorem componen we write it only for  $\widehat{\tau}_{iskt}$ for sake of notational simplicity. For the same reason, we do not distinguish the coefficients associated to the fixed effects between Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), even if they are specific to the type of transport costs considered (e.g., the series of $\gamma_t$ differs from one estimation to the other).}
%$^{,}$\footnote{We use the transport costs series obtained at the 3-digit classification level.}
